---
title: "Multithreshold approach with varying number of functions and species"
output:
  html_document: default
---

This script produces:

+ Figure 4

This script sets up the simulations to show the effect of including a varying number of functions and (separately) a varying number of species on the slope pattern produced by the multithreshold approach. 

For the **variable number of function simulation** we hold species richness constant at `specnum`. 

We then define a set number of functions of size `funcnum` from which we draw all possible (but max 50) subsets of variable size (3 subsets-sizes total). For each subset of functions we calculate the multithreshold approach. 

For the **variable number of species simulation** we hold the number of functions constant at `funcnum` but calculate the multithreshold approach for the full species range and two smaller subsets.  


```{r, echo = FALSE, warning=FALSE, message=FALSE, "load packages"}
library(multifunc)
library(dplyr)
library(tidyr)
library(ggplot2)
library(cowplot)

source("Multifunc_simulations_functions.R")
```

# simulate full diversity experiment

One can set the same parameters as in most other simulations:

+ `distribution` : the distribution function. The names of the parameters must be changed accordingly in `FunctionValue()`
+ `specnum` : the (maximum) number of species
+ `funcnum` : the (maximum) number of functions 
+ `method` : the method to use (with or without complementarity)

Additional parameters for `method = comp`:

+ `CF` : maximum complementarity factor 
+ `compfunc` : which functions should experience complementarity (`all` or any combination of `func.names`)
+ `r` : the *growthrate* of the complementarity factor


```{r}
set.seed(777)

specnum <- 15
funcnum <- 9

distribution = "runif"

FuncMat <- FunctionValue(specnum,funcnum, distribution, min = 0, max = 1)

func.names <- as.character( unique( FuncMat$Functions))
spec.names <- as.character( unique( FuncMat$Species))

#maxrep <- choose(specnum, floor(specnum/2))
maxrep <- 200 #using the full replications is prohibitive

SpecMat <- SpeciesMatrix(specnum = specnum, maxrep = maxrep)

method = "av"

AvFunc <- AverageFunction(SpecMat, FuncMat,
                          method = method, 
                          CF = CF, 
                          compfunc = compfunc,
                          r = r)

# standardize functions
AvFunc_func <- AvFunc %>% 
  mutate_at(vars(one_of(func.names)), function(x) {(x - min(x)) / (max(x) - min(x))})
```

### Variable number of function simulation

```{r}
# empty dataframe to store results
RES_func <- data.frame(thresholds = numeric(), 
                   Estimate = numeric(), 
                   nfunc = numeric(),
                   func_comb = numeric())

#loop over chosen subsets of all function of varying size
for (i in c(ceiling(funcnum/3), 2*ceiling(funcnum/3), funcnum)) { 

  # all poosibel combination of i out of funcnum functions
  func_comb <- combn(func.names, i)
  
  # sample 50 random function combinations if more than 50 possible combinations
  if(ncol(func_comb) > 50) {
    func_comb <- func_comb[, sample(c(1:ncol(func_comb)), 50)]
  }

  #loop over all function combinations of size i
  for ( k  in seq_len(ncol(func_comb))) { 
    
    # number of functions above threshold
    mixedThresh <- getFuncsMaxed(AvFunc, func_comb[ ,k], threshmin=0.05,
                                 threshmax=0.99, prepend=c("Richness"), maxN=1)
    # slopes  
    mixedLinearSlopes<-getCoefTab(funcMaxed ~ Richness, fun = lm, 
                                  data=mixedThresh,coefVar="Richness")
    
    colnames(mixedLinearSlopes) <- c("thresholds", "Estimate",
                                     "Std. Error", "t value", "Pr(>|t|)")
    
    temp <- mixedLinearSlopes %>% 
      select(thresholds, Estimate) %>% 
      mutate(nfunc = i) %>% 
      mutate(func_comb = k)
    
    RES_func <- rbind(RES_func, temp)
  }
  }

```

### Plot with variable number of functions

```{r}

FUNC <- RES_func %>% 
  group_by(thresholds, nfunc) %>% 
  summarise(mean_Estimate = mean(Estimate),
            CI_high = mean(Estimate) + 1.96 * (sd(Estimate)/sqrt(n())),
            CI_low = mean(Estimate) - 1.96 * (sd(Estimate)/sqrt(n()))) %>% 
ggplot(., aes(x=thresholds*100, y=mean_Estimate), size = 0.5, alpha = 0.3)+
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high, fill = as.factor(nfunc)), colour = NA, alpha = 0.4)+
  geom_line( aes(colour = as.factor(nfunc)), lwd = 0.8) +
  ylab("Slope estimate") + xlab("Threshold (%)") +
  geom_abline(intercept=0, slope=0, lwd=0.5, linetype=2) + 
  theme_bw(base_size=15)+
  scale_fill_brewer(guide = FALSE, palette = "Set1")+
  scale_color_brewer(guide = guide_legend(title = paste("Number of functions", paste("(", specnum, " species)", sep = ""), sep = "\n")),
                     palette = "Set1")+
  theme_classic()+
  theme(legend.position = "bottom")+
  scale_y_continuous(limits = c(-0.45, 0.45)) 
  
FUNC
```

### Variable number of species simulation

```{r}

# empty dataframe to store results
RES_spec <- data.frame(thresholds = numeric(), 
                   Estimate = numeric(),
                   `Std. Error` = numeric(), 
                   nspec = numeric())

#loop over three subsets of specnum
for (i in c(6, 9, specnum)) { 

  #subset for number of species to include
  AvFunc_spec <- filter(AvFunc, Richness %in% 1:i) 
  
  # number of functions above threshold
  mixedThresh <- getFuncsMaxed(AvFunc_spec, func.names,
                               threshmin=0.05, threshmax=0.99,
                               prepend=c("Richness"), maxN=1)

  #slopes
  mixedLinearSlopes<-getCoefTab(funcMaxed ~ Richness, fun = lm,
                                data=mixedThresh, coefVar="Richness")
    
  colnames(mixedLinearSlopes) <- c("thresholds", "Estimate",
                                   "Std. Error", "t value", "Pr(>|t|)")
    
  temp <- mixedLinearSlopes %>% 
      select(thresholds, Estimate, `Std. Error`) %>% 
      mutate(nspec = i)
    
    RES_spec <- rbind(RES_spec, temp)
  }

```

### Plot with variable number of species

```{r}
#plot
SPEC <- RES_spec %>% 
  group_by(thresholds, nspec) %>% 
  mutate(CI_high = Estimate + 1.96 * `Std. Error`,
            CI_low = Estimate - 1.96 * `Std. Error`) %>% 
ggplot(., aes(x=thresholds*100, y=Estimate), size = 0.5, alpha = 0.3)+
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high, fill = as.factor(nspec)), colour = NA, alpha = 0.4)+
  geom_line( aes(colour = as.factor(nspec)), lwd = 0.8) +
  ylab("Slope estimate") + xlab("Threshold (%)") +
  geom_abline(intercept=0, slope=0, lwd=0.5, linetype=2) + 
  theme_bw(base_size=15)+
  scale_fill_brewer(guide = FALSE, palette = "Set1")+
  scale_color_brewer(guide = guide_legend(title = paste("Number of species", paste("(", funcnum, " functions)", sep = ""), sep = "\n")),
                     palette = "Set1")+
  theme_classic()+
  theme(legend.position = "bottom")+
  scale_y_continuous(limits = c(-0.45, 0.45)) 
  
SPEC
```

### Arrange plots
```{r}
# arrange plots and save
plot <- plot_grid(FUNC, SPEC, nrow = 1,
                      labels = c("a", "b"),
                      hjust = -7,
                      vjust = 3,
                      label_size = 15)

plot
ggsave("Figure_4.pdf", plot, height = 4, width = 8)

```

