---
Title: "sensibility of multithreshold approach"
output:
  html_document: default
---

This script produces:

+ Figure 6
+ Figure S1

This script sets up the simulations to show the effect of standardization on the multithreshold approach. It includes 4 different scenarios.

+ `Single_dist`: 
All function values are drawn from a single distribution. **Choosing the pre-set values produces Figure S1**

+ `Multi_dist`:
Different parameters can be specified for each function (which are supplied to the same distribution).

+ `Biodepth_data`:
The monoculture data from the BIODEPTH experiment (included in the `multifunc` package) are used to simulate the null-expectation for the multithreshold approach. **Choosing Germany produces Figure 6**

+ `Biodepth_sim
The mean and standard deviation of the function from the monoculture data from the BIODEPTH experiment (included in the `multifunc` package) are used as distribution values for a normal distribution. Special case of `Multi_dist`.


```{r, echo = FALSE, message = FALSE, warning=FALSE, "packages"}
library(multifunc) #make sure that you don't have dplyr allready loaded otherwise the code will fail because of name conflicts
library(dplyr)
library(tidyr)
library(ggplot2)
library(gridExtra)
library(cowplot)
library(RColorBrewer)

source("Multifunc_simulations_functions.R")
```

### Choose simulation

+ choose the simulation that you want to perform by setting the value to `TRUE`
```{r, "choose country"}
# choose model (by setting vlaue to TRUE)

Single_dist <- FALSE #produces Figure S1
Multi_dist <- FALSE
Biodepth_data <- TRUE #produces Figure 6
Biodepth_sim <- FALSE


if(sum(Single_dist,Multi_dist,Biodepth_data, Biodepth_sim) != 1){stop("please choose one (and only one) scenario by setting the value to TRUE")}
```

### Choose country (if `Biodepth_data` or `Biodepth_sim`)

```{r, `choose simulation`}
if(sum(Biodepth_data, Biodepth_sim) == 1) {
  
  data("all_biodepth")
  
#choose Location (one off "Germany" "Portugal" "Switzerland" "Greece" "Ireland" "Sweden" "Sheffield" "Silwood" )
  
  Location <- "Germany"
  
  }

```

### Scenario `Single_dist`

All function values are drawn from a single distribution. **Choosing the pre-set values produces Figure S1**

You can set the same parameters as in most other simulations:

+ `distribution` : the distribution function. The names of the parameters must be changed accordingly in `FunctionValue()`
+ `specnum` : the number of species
+ `funcnum` : the number of functions 
+ `method` : the method to use (with or without complementarity)

Additional parameters for `method = comp`:

+ `CF` : maximum complementarity factor 
+ `compfunc` : which functions should experience complementarity (`all` or any combination of `func.names`)
+ `r` : the 'growth-rate' of the complementarity factor

One can choose a random seed by commenting out `set.seed` or you can specify your own seed. 

```{r, "Single_dist"}

if(Single_dist) {
  
set.seed(999)

specnum <- 15
funcnum <- 9

distribution = "rnorm" #choose your distribution (one off ?Distributions)

FuncMat <- FunctionValue(specnum,funcnum, distribution, mean = 30, sd = 6)  

func.names <- as.character( unique( FuncMat$Functions))

spec.names <- as.character( unique( FuncMat$Species))

SpecMat <- SpeciesMatrix(specnum = specnum, maxrep = 200)

method = "av" #av or comp

AvFunc <- AverageFunction(SpecMat, FuncMat,
                          method = method, 
                          CF = CF, 
                          compfunc = compfunc,
                          r = r)

}
  
```

### Scenario `Multi_dist`

Different parameters can be specified for each function (which are supplied to the same distribution function). The parameters (`par1` and `par2`, the second one is not need for all distributions) must be specified as list of length `funcnum`. Otherwise as `Single_dist`.

```{r, "Multi_dist"}

if(Multi_dist) {
  
set.seed(999)

specnum <- 15
funcnum <- 9

distribution = "rnorm" #choose your distribution (one off ?Distributions)


FuncList <- list()

# choose parameters
par1 <- seq(50, 100, len = funcnum) #parameter values for functions 1:nfunc (first parameter)
par2 <- seq(15,3, len = funcnum) #parameter values for functions 1:nfunc (second parameter - if needed)

for (i in 1:funcnum){
var <- paste("FuncMat", i, sep = "_")

# you have to name the parameters according to your chosen distribution
FuncList[[var]] <- FunctionValue(specnum,funcnum, distribution, mean = par1[[i]], sd = par2[[i]])  
}

FuncMat <- FuncList[[1]]

func.names <- as.character( unique( FuncMat$Functions))

for(n in seq_along(func.names)){
  FuncMat[FuncMat$Functions == func.names[n], ]$Funcval <- FuncList[[n]][FuncList[[n]]$Functions == func.names[n], ]$Funcval
}

spec.names <- as.character( unique( FuncMat$Species))

SpecMat <- SpeciesMatrix(specnum = specnum, maxrep = 200)

method = "av"

AvFunc <- AverageFunction(SpecMat, FuncMat,
                          method = method, 
                          CF = CF, 
                          compfunc = compfunc,
                          r = r)
}
```


### Scenario `Biodepth_data`

The monoculture data from the BIODEPTH experiment (included in the `multifunc` package) are taken to simulate the null-expectation for the multithreshold approach. **Choosing Germany produces Figure 5**

you can choose to simulate complementarity for some or all functions

```{r, "Biodepth_data"}

if(Biodepth_data) {

  set.seed(999)
  
biodepth_func <- 
  all_biodepth %>% 
  gather("species", "p_a", 26:ncol(.)) %>% 
  filter(p_a == 1) %>% 
  filter(Diversity == 1) %>% 
  filter(location == Location) %>% 
  select(species, 15:25) %>% 
  select_if(function(col) sum(is.na(col)) <= floor(0.2*length(col))) %>% #exclude functions with > 20% NAs
  na.omit() %>% #exclude species with missing function values
  group_by(species) %>% 
  summarise_all(function(x) mean(x, na.rm = T)) %>% 
  rename("Species" = species)

specnum <- nrow(biodepth_func)
spec.names <- biodepth_func$Species


funcnum <- ncol(biodepth_func)-1


FuncMat <- gather(biodepth_func, Functions, Funcval, -Species)
func.names <- as.character( unique( FuncMat$Functions))

SpecMat <- SpeciesMatrix(specnum, maxrep = 511)
dimnames(SpecMat)[[2]] <- spec.names

method = "av"

AvFunc <- AverageFunction(SpecMat, FuncMat,
                          method = method, 
                          CF = CF, 
                          compfunc = compfunc,
                          r = r)

}

```

### Scenario `Biodepth_sim`

The mean and standard deviation of the function from the monoculture data from the BIODEPTH experiment (included in the `multifunc` package) are used as distribution values for a normal distribution. Special case of `Multi_dist`.

```{r, "Biodepth_sim"}

if(Biodepth_sim) {

  set.seed(999)

biodepth_func <- 
  all_biodepth %>% 
  gather("species", "p_a", 26:ncol(.)) %>% 
  filter(p_a == 1) %>% 
  filter(Diversity == 1) %>% 
  filter(location == Location) %T>% 
  assign('Loc_sub', ., envir = .GlobalEnv) %>% 
  select(species, 15:25) %>% 
  select_if(function(col) sum(is.na(col)) <= floor(0.2*length(col))) %>% #exclude functions with > 20% NAs
  gather(Function, value, -species) %>% 
  group_by(Function) %>% 
  summarise(mean_func = mean(value, na.rm = T),
            sd_func = sd(value, na.rm = T))

spec.names <- unique(Loc_sub$species)
specnum <- length(spec.names)

funcnum <- nrow(biodepth_func)

distribution = "rnorm" 

# if function values should be drawn form distributiuon with differnt means

FuncList <- list()

par1 <- biodepth_func$mean_func #parameter values for functions 1:nfunc (first parameter)
par2 <- biodepth_func$sd_func #parameter values for functions 1:nfunc (second parameter - if needed)

for (i in 1:funcnum){
var <- paste("FuncMat", i, sep = "_")

# you have to name the parameters according to your chosen distribution
FuncList[[var]] <- FunctionValue(specnum,funcnum, distribution, mean = par1[[i]], sd = par2[[i]])  
}

FuncMat <- FuncList[[1]]

func.names <- as.character( unique( FuncMat$Functions))

for(n in seq_along(func.names)){
  FuncMat[FuncMat$Functions == func.names[n], ]$Funcval <- FuncList[[n]][FuncList[[n]]$Functions == func.names[n], ]$Funcval
}

spec.names <- as.character( unique( FuncMat$Species))

SpecMat <- SpeciesMatrix(specnum = specnum, maxrep = 200)

method = "av"

AvFunc <- AverageFunction(SpecMat, FuncMat,
                          method = method, 
                          CF = CF, 
                          compfunc = compfunc,
                          r = r)
}

```

### Slope simualtion

Standardize the function values with different methods (between 0 and 1 and by the maximum) and calculate the multithreshold approach for both scenarios.

```{r, "Slope simualtion"}

# standardize functions by maximum
AvFunc_max <- AvFunc %>% 
  mutate_at(vars(one_of(func.names)), function(x) {x / max(x)})

# standardize functions between 0 and 1
AvFunc_unit <- AvFunc %>% 
   mutate_at(vars(one_of(func.names)), function(x) {(x - min(x)) / (max(x) - min(x))})

# store both dataset in list
Av_list <- list(AvFunc_max = AvFunc_max,
                AvFunc_unit = AvFunc_unit)

# empty dataframe to store results
RES_stand <- data.frame(thresholds = numeric(), 
                   Estimate = numeric(),
                   `Std. Error` = numeric(), 
                   stand_method = character())


for (i in seq_along(Av_list)) { #loop over differnt standardizations
  
  mixedThresh <- getFuncsMaxed(Av_list[[i]], func.names, threshmin=0.05, threshmax=0.99, prepend=c("Richness"), maxN=1)
  
  mixedLinearSlopes<-getCoefTab(funcMaxed ~ Richness, fun = lm,  data=mixedThresh, 
                               coefVar="Richness")
    
  colnames(mixedLinearSlopes) <- c("thresholds", "Estimate",  "Std. Error", "t value", "Pr(>|t|)")
    
  temp <- mixedLinearSlopes %>% 
      select(thresholds, Estimate, `Std. Error`) %>% 
      mutate(stand_method = names(Av_list)[i])
    
    RES_stand <- rbind(RES_stand, temp)
  }

```

### Plot slope patterns

```{r, "Plot slope patterns"}

STAND <- RES_stand %>% 
  group_by(thresholds, stand_method) %>% 
  mutate(CI_high = Estimate + 1.96 * `Std. Error`,
            CI_low = Estimate - 1.96 * `Std. Error`) %>% 
ggplot(., aes(x=thresholds*100, y=Estimate), size = 0.5, alpha = 0.3)+
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high, fill = as.factor(stand_method)), colour = NA, alpha = 0.4)+
  geom_line( aes(colour = as.factor(stand_method)), lwd = 0.8) +
  ylab("Slope estimate") + xlab("Threshold (%)") +
  geom_abline(intercept=0, slope=0, lwd=0.5, linetype=2) + 
  theme_bw(base_size=15)+
  scale_fill_brewer(guide = FALSE, palette = "Set1")+
  scale_color_brewer(guide = guide_legend(title =""), palette = "Set1", labels = c("scaled by max", "scaled [0,1]"))+
  theme_classic()+
  theme(legend.position = c(.2,0.3),  legend.background = element_rect(fill = "transparent", colour = "transparent"))
  
  
STAND
```

### Plot distribution of raw function values

```{r, "raw func"}

get_colours <- colorRampPalette(brewer.pal(9, "Greys"))

Hist_orig <- AvFunc %>% 
  select(one_of(func.names)) %>% 
  gather(Func, value) %>% 
  ggplot(.)+
  geom_density(aes(x = value,y =..density.., colour = Func), position="identity")+
  scale_colour_manual( values = get_colours(funcnum))+
  theme_classic()+
  theme(legend.position = "none") +
  labs(x = "Simulated function values", y = "Density")

Hist_orig

```

### Plot distribution of standardized function values (between 0 and 1)

```{r, "unit_func"}

Hist_unit <- AvFunc_unit %>% 
  select(one_of(func.names)) %>% 
  gather(Func, value) 


  maxD <- Hist_unit %>% 
  group_by(Func) %>% 
  summarize(maxD = max(density(value)$y)) %>% 
  ungroup %>% 
  summarize(maxD = max(maxD)) %>% as.numeric()

get_colours <- colorRampPalette(brewer.pal(9, "Blues"))
    
Hist_unit_gg <- ggplot(Hist_unit)+
  geom_density(aes(x = value,y =..density.., colour = Func), position="identity")+
  geom_vline(aes(xintercept = mean(value)), colour = "black", linetype = "dashed")+
  annotate("label", x = 0.8, y = 0.8*maxD,
           label = "f(x) == frac(x - plain(min)(x), plain(max)(x) - plain(min)(x))", 
           parse=T, size = 2.5, alpha = 0.8)+
  geom_vline(aes(xintercept = mean(value)), colour = "black", linetype = "dashed")+
  scale_colour_manual(values = get_colours(funcnum))+
  theme_classic()+
  theme(legend.position = "none")+
  labs(x = "Scaled between [0,1]", y = "Density")

Hist_unit_gg

```


### Plot distribution of standardized function values (by maximum)

```{r, "max_func"}

Hist_max <- AvFunc_max %>% 
  select(one_of(func.names)) %>% 
  gather(Func, value) 

maxD <- Hist_max %>% 
  group_by(Func) %>% 
  summarize(maxD = max(density(value)$y)) %>% 
  ungroup %>% 
  summarize(maxD = max(maxD)) %>% as.numeric()

get_colours <- colorRampPalette(brewer.pal(9, "Reds"))
  
Hist_max_gg <- ggplot(Hist_max)+
  geom_density(aes(x = value,y =..density.., colour = Func), position="identity")+
  geom_vline(aes(xintercept = mean(value)), colour = "black", linetype = "dashed")+
  annotate("label", x = 0.3, y = 0.8*maxD, 
           label = "f(x) == frac(x , plain(max)(x))", 
           parse=T, size = 2.5)+
  scale_colour_manual(values = get_colours(funcnum))+
  scale_x_continuous(limits = c(0,1))+
  theme_classic()+
  theme(legend.position = "none")+
  labs(x = "Scaled by maximum", y = "Density")

Hist_max_gg
```

### Arrange, draw and save the plots

```{r, fig.height= 6, fig.width=10}
if(sum(Biodepth_data + Biodepth_sim) ==1) {
  plot <- 
  ggdraw() +
  draw_plot(add_sub(STAND, "a", 0.02,8), 0, .5, 1, .5) +
  draw_plot(add_sub(Hist_unit_gg,"b", 0.05, 8), 0, 0, 0.5, 0.5)+
  draw_plot(add_sub(Hist_max_gg,"c", 0.05, 8), 0.5, 0, 0.5, 0.5)
} else {
  plot <- 
  ggdraw() +
  draw_plot(add_sub(STAND, "a", 0.02,8), 0, .5, 1, .5) +
  draw_plot(add_sub(Hist_orig,"b", 0.05,8), 0, 0, 1/3, 0.5) +
  draw_plot(add_sub(Hist_unit_gg,"c", 0.05, 8), 1/3, 0, 1/3, 0.5)+
  draw_plot(add_sub(Hist_max_gg,"d", 0.05, 8), 2/3, 0, 1/3, 0.5)
}

plot

if(Biodepth_data) {
  ggsave("Figure_6.pdf", plot, height = 6, width = 10)
}

if(Single_dist) {
  ggsave("Figure_S1.pdf", plot, height = 6, width = 10)
}

```

