---
title: "Average and multithreshold approach with and w/o complementarity"
output: html_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

```

This script produces:

+ Figure 2


```{r, "load packages"}

library(ggplot2)
library(Cairo)
library(grid)
library(cowplot)
library(dplyr)
library(tidyr)
library(corrplot)
library(magrittr)

source("Multifunc_simulations_functions.R")
```


### Simulate diversity experiment

In this script we set up the simulations for the null expectation of the three most common multifunctionality metrics: 

+ the average multifunctionality approach
+ the single threshold approach
+ the multiple threshold approach

For this (and all other) simulations we set a number of species with the `specnum` parameter and a number of functions these species perform with the `funcnum` parameter. The function values for each species and functions are then drawn at at random from a chosen distribution set by the `distribution` parameter. See `?Distributions` for possible choices and the required parameters. The default is a uniform distribution in the interval [0,1]. 

With the species and function values defined, we simulate a diversity experiment where 

+ we have 1 : `specnum` richness levels
+ we have all possible combinations at each level (or a defined maximum of unique species combinations- set with `maxrep`; recommended for `specnum` > 12)
+ we need to choose a method for computing a mixture function value. 

The possible methods are:

+ `av` for average. Here the mixture values at each diversity level are the unweighted single - species function values. So simply the mean (or average) value of the community.

+ `comp` for complementarity. Here one, several or all functions are multiplied with a complementary factor before the average is calculated. The complementary factor is always a saturating  function of the richness level of the form:

$$ compfac = CF * ( 1 - ( 1 - \frac{1}{CF} ) * exp{ ( 1 - Richness ^ r )} $$

  + `CF` : maximum value for the complementary factors as $S \to \inf$
  + `r` *growthrate* i.e. the rate at which `compfac` grows to `CF`
  + `compfunc` : the functions subjected to complementarity Either "all" or a subset of `func.names`
  
To make the analysis reproducible we can set a seed with `set.seed. 

### Draw function values for all species

`FunctionValue` takes three main arguments, `specnum`, `funcnum` and `distribution` and returns a data-frame with function values for all species and functions. 

** If you change the distribution you have to change the distribution paramters in `FunctionValue`**

```{r, "funcmat"}
# set seed to replicate analysis:
set.seed(999)

# number of species
specnum <- 15

# number of functions
funcnum <- 9

# distribution from which to draw function values
distribution = "runif"

#DF <- data.frame(seed = rep(NA, 100), mean = NA)


# create function matrix
FuncMat <- FunctionValue(specnum,funcnum, distribution, min = 0, max = 1)

```

### Plot function values and function correlations
```{r, "corrplot"}

FuncMat_plot <- 
  FuncMat %>% 
  group_by(Functions) %>% 
  mutate(Funcval = (Funcval - min(Funcval)) / (max(Funcval) - min(Funcval))) %>% 
  spread(Functions, Funcval) %>% 
  select(-Species)
  
  
FuncMat_plot %>% 
  mutate(av = rowMeans(.)) %>%
  as.matrix() %>% 
  set_rownames(unique(FuncMat$Species)) %>%
  corrplot( is.corr = F, tl.col = "black", cl.ratio = 0.2, 
           cl.length = 11, method="square")

FuncMat_plot %>% cor %>% 
  corrplot(type = "lower", tl.col = "black", cl.ratio = 0.2, 
           cl.length = 11, number.cex = 0.6, addCoef.col = "#323232", diag = F, method="ellipse")

```


## Scenario 1 - no diveristy effect on any single function

### Simulate diversity experiment

+ `SpeciesMatrix` takes two arguments, `specnum`and `maxrep` and generates a *plot x species* matrix where each row is a *plot* and each column a *species*. Cell values are 0 or 1 and indicate which species are present in which plot. If `maxrep` is higher than `choose(specnum, floor(specnum/2))` (i.e. choose(15,7) = 6435 possible combinations at richness level 7 for 15 species) all unique species combinations are simulated at each richness level. If `maxrep` is set to a lower number, a maximum of `maxrep` unique combinations are simulated for each richness level. 

+ `AverageFunction` takes the out put of `SpeciesMatrix` and `FunctionValue`, i.e. the plot x species matrix and the species - function data-frame simulated above, to calculate plot-wise functioning for all plots, given the chosen method (`av`or `comp`)

```{r, "Avfunc"}
# code takes ~5 min to run for 15 species. Set to lower value for higher number of species! 

maxrep <- choose(specnum, floor(specnum/2))
#maxrep <- 500

# simulate plot x species matrix
SpecMat <- SpeciesMatrix(specnum = specnum, maxrep = maxrep)

# select method and additional parameters if appropriate by setting the `if` statement to `TRUE`
if (TRUE) {method = "av"}

if (FALSE) {method = "comp"
            CF = 3
            compfunc = c("F 1", "F 6")
            r = 0.25}


# Average function
AvFunc <- AverageFunction(SpecMat, FuncMat,
                          method = method, 
                          CF = CF, 
                          compfunc = compfunc,
                          r = r)
```

### Standardize function values

There are two main ways to standardize the function values:

+ between 0 and 1 with $x_{stand} = \frac{x - min(x)}{max(x) - min(x)}$

+ by the maximum with $x_{stand} = \frac{x}{max(x)}$

```{r, "standardize"}
# extract function names
func.names <- as.character( unique( FuncMat$Functions))

# standardize by maximum
 AvFunc[,func.names] <- apply(AvFunc[,func.names], 2, function(x) {x/max(x)})

# alternative standardize between 0 and 1
# AvFunc[,func.names] <- apply(AvFunc[,func.names], 2, function(x) {(x - min(x)) / (max(x) - min(x))})

```

### Plot single function values
```{r, "single_func"}
AvFunc_long <- gather(AvFunc[, -c(1:specnum)], Function, FuncVal, -Richness) 

SingleFunc <- AvFunc_long %>%
  mutate(FuncVal = round(FuncVal,3)) %>% 
  group_by(Richness, Function, FuncVal) %>% 
  summarize(n = n()) %>% 
  ggplot(., aes(x=Richness, y=FuncVal, colour = n))+
  geom_point( size = 1, position = position_jitter(width = 0.1))+
  facet_wrap(~Function)+
  theme_bw(base_size=15)+
  stat_smooth(data = AvFunc_long, aes(x = Richness, y = FuncVal), method="lm", colour = "#de2d26") +
  xlab("\nSpecies richness") +
  ylab("Value of function\n") +
  theme(panel.grid = element_blank(), legend.position = "none")

SingleFunc

```

### Average approach for multifunctionality

+ Multifunctionality is calculated as the average value of the standardized single functions.

```{r, "multifunc_av"}

# add averaged multifunctional index
AvFunc$meanFunction <- rowMeans(AvFunc[,func.names])

# plot
AverageFunc <- 
  AvFunc %>%
  select(Richness, meanFunction) %>% 
  mutate(meanFunction = round(meanFunction,3)) %>% 
  group_by(Richness, meanFunction) %>% 
  summarize(n = n()) %>% 
  ggplot(., aes(x=Richness, y=meanFunction, colour = n))+
  geom_point(size = 2, position = position_jitter(width = 0.1))+
    theme_bw(base_size=15)+
    stat_smooth(data = AvFunc, aes(x = Richness, y = meanFunction), method="lm", colour = "#de2d26") +
    xlab("\nSpecies richness") +
    ylab("Average value of standardized functions\n")+
  theme(legend.position = "none")+
   # scale_y_continuous(limits = c(0.25,0.75))+
    scale_x_continuous(breaks = seq(1,15,2))

AverageFunc

```

### Single-threshold approach for multifunctionality

+ Multifunctionality is calculated as the number of functions that are sustained above a pre-defined threshold - calculated as a percentage of the maximum observed function value across all plots.

```{r, "single_thresh"}

mixedThresh <- getFuncsMaxed(AvFunc, func.names, threshmin=0.05, threshmax=0.99, 
                           prepend=c("Richness"), maxN=1)


singleThresh <- filter(mixedThresh, as.character(thresh) %in% as.character(seq(0,1,0.1))) %>% 
  mutate(prct = paste(thresholds * 100, "%")) %>% 
  select(Richness, funcMaxed, prct) %>% 
  group_by(Richness, funcMaxed, prct) %>% 
  summarise(n = n()) %>% 
  ggplot(., aes(x = Richness, y = funcMaxed, fill = n))+
  geom_point(shape = 21, size = 2.5, colour = "black", stroke = 0.2)+
  stat_smooth(method = "lm", colour = "#de2d26", se = F)+
  facet_wrap(~prct)+
  labs(x = "Species richness", y = "Number of functions ≥ threshold")+
  theme_bw(base_size=15)+
  scale_y_continuous(limits = c(0,10), breaks = seq(0,10,3))+
  scale_x_continuous(breaks = seq(1,15,2))+
  theme(legend.position = "none")

singleThresh
```

### Multi-threshold approach for multifunctionality

+ The slopes shown in the previous figure are plotted against threshold values over the whole range of possible thresholds.

```{r, "multithresh"}
mixedLinearSlopes <- getCoefTab(funcMaxed ~ Richness, fun = lm,  data=mixedThresh, 
                               coefVar="Richness")

p <- SlopeSummary(mixedLinearSlopes)

multiTresh <- ggplot(mixedLinearSlopes, aes(x=thresholds)) +
  geom_ribbon(fill="#deebf7", aes(x=thresholds*100, ymin=Estimate- 1.96*mixedLinearSlopes[["Std. Error"]],                              ymax=Estimate+1.96*mixedLinearSlopes[["Std. Error"]])) + 
  geom_point(aes(x=thresholds*100, y=Estimate), colour = "#132B43") +
  annotate("segment", 
           x = p$Estimate_sign_change*100 + 7,
           xend = p$Estimate_sign_change*100,
           y = 0.25*max(mixedLinearSlopes$Estimate),
           yend = 0)+
  annotate("label",
           x = p$Estimate_sign_change*100 + 7,
           y = 0.25*max(mixedLinearSlopes$Estimate),
           label = paste("p: ", round(p$Estimate_sign_change*100), "%", sep = ""))+
  ylab("Slope") + xlab("Threshold (%)") +
  geom_abline(intercept=0, slope=0, lwd=1, linetype=2) + 
  theme_bw(base_size=15)+
  scale_y_continuous(limits = c(-0.3, 0.3))
  

multiTresh 
```



## Scenario 2 - complementarity effect on 2 functions

Exact same code as above, but with the `method = "comp"` chosen to simulate complementary for two functions. 

### Simulate diversity experiment


```{r, "Avfunc_comp"}
# code takes ~5 min to run for 15 species. Set to lower value for higher number of species! 

#maxrep <- choose(specnum, floor(specnum/2))
#maxrep <- 500

# simulate plot x species matrix (If not allready done for Scenario 1)
#SpecMat <- SpeciesMatrix(specnum = specnum, maxrep = maxrep)

# select method and additional parameters if appropriate by setting the `if` statement to `TRUE`
if (FALSE) {method = "av"}

if (TRUE) {method = "comp"
            CF = 3
            compfunc = c("F 1", "F 6")
            r = 0.25}


# Average function
AvFunc <- AverageFunction(SpecMat, FuncMat,
                          method = method, 
                          CF = CF, 
                          compfunc = compfunc,
                          r = r)
```

### Standardize function values

```{r, "standardize_comp"}
# extract function names
func.names <- as.character( unique( FuncMat$Functions))

# standardize by maximum
 AvFunc[,func.names] <- apply(AvFunc[,func.names], 2, function(x) {x/max(x)})

# alternative standardize between 0 and 1
# AvFunc[,func.names] <- apply(AvFunc[,func.names], 2, function(x) {(x - min(x)) / (max(x) - min(x))})

```

### Plot single function values
```{r, "single_func_comp"}
AvFunc_long <- gather(AvFunc[, -c(1:specnum)], Function, FuncVal, -Richness) 

SingleFunc_comp <- AvFunc_long %>%
  mutate(FuncVal = round(FuncVal,3)) %>% 
  group_by(Richness, Function, FuncVal) %>% 
  summarize(n = n()) %>% 
  ggplot(., aes(x=Richness, y=FuncVal, colour = n))+
  geom_point( size = 1, position = position_jitter(width = 0.1))+
  facet_wrap(~Function)+
  theme_bw(base_size=15)+
  stat_smooth(data = AvFunc_long, aes(x = Richness, y = FuncVal), method="lm", colour = "#de2d26") +
  xlab("\nSpecies richness") +
  ylab("Value of function\n") +
  theme(panel.grid = element_blank(), legend.position = "none")

SingleFunc_comp

```

### Average approach for multifunctionality
```{r, "multifunc_av_comp"}

# add averaged multifunctional index
AvFunc$meanFunction <- rowMeans(AvFunc[,func.names])

#plot it
AverageFunc_comp <- 
  AvFunc %>%
  select(Richness, meanFunction) %>% 
  mutate(meanFunction = round(meanFunction,3)) %>% 
  group_by(Richness, meanFunction) %>% 
  summarize(n = n()) %>% 
  ggplot(., aes(x=Richness, y=meanFunction, colour = n))+
  geom_point(size = 2, position = position_jitter(width = 0.1))+
    theme_bw(base_size=15)+
    stat_smooth(data = AvFunc, aes(x = Richness, y = meanFunction), method="lm", colour = "#de2d26") +
    xlab("\nSpecies richness") +
    ylab("Average value of standardized functions\n")+
  theme(legend.position = "none")+
   # scale_y_continuous(limits = c(0.25,0.75))+
    scale_x_continuous(breaks = seq(1,15,2))

AverageFunc_comp

```

### Single-threshold approach for multifunctionality
```{r, "single_thresh_comp"}

mixedThresh <- getFuncsMaxed(AvFunc, func.names, threshmin=0.05, threshmax=0.99, 
                           prepend=c("Richness"), maxN=1)


singleThresh_comp <- filter(mixedThresh, as.character(thresh) %in% as.character(seq(0,1,0.1))) %>% 
  mutate(prct = paste(thresholds * 100, "%")) %>% 
  select(Richness, funcMaxed, prct) %>% 
  group_by(Richness, funcMaxed, prct) %>% 
  summarise(n = n()) %>% 
  ggplot(., aes(x = Richness, y = funcMaxed, fill = n))+
  geom_point(shape = 21, size = 2.5, colour = "black", stroke = 0.2)+
  stat_smooth(method = "lm", colour = "#de2d26", se = F)+
  facet_wrap(~prct)+
  labs(x = "Species richness", y = "Number of functions ≥ threshold")+
  theme_bw(base_size=15)+
  scale_y_continuous(limits = c(0,10), breaks = seq(0,10,3))+
  scale_x_continuous(breaks = seq(1,15,2))+
  theme(legend.position = "none")

singleThresh_comp
```

### Multi-threshold approach for multifunctionality
```{r, "multithresh_comp"}
mixedLinearSlopes<-getCoefTab(funcMaxed ~ Richness, fun = lm,  data=mixedThresh, 
                               coefVar="Richness")

colnames(mixedLinearSlopes) <- c("thresholds", "Estimate",  "Std. Error", "t value", "Pr(>|t|)")

p <- SlopeSummary(mixedLinearSlopes)

multiTresh_comp <- ggplot(mixedLinearSlopes, aes(x=thresholds)) +
  geom_ribbon(fill="#deebf7", aes(x=thresholds*100, ymin=Estimate- 1.96*mixedLinearSlopes[["Std. Error"]],                              ymax=Estimate+1.96*mixedLinearSlopes[["Std. Error"]])) + 
  geom_point(aes(x=thresholds*100, y=Estimate), colour = "#132B43") +
  annotate("segment", 
           x = p$Estimate_sign_change*100 + 7,
           xend = p$Estimate_sign_change*100,
           y = 0.25*max(mixedLinearSlopes$Estimate),
           yend = 0)+
  annotate("label",
           x = p$Estimate_sign_change*100 + 7,
           y = 0.25*max(mixedLinearSlopes$Estimate),
           label = paste("p: ", round(p$Estimate_sign_change*100), "%", sep = ""))+
  ylab("Slope") + xlab("Threshold (%)") +
  geom_abline(intercept=0, slope=0, lwd=1, linetype=2) + 
  theme_bw(base_size=15)+
  scale_y_continuous(limits = c(-0.3, 0.3))
  

multiTresh_comp 
```

### Figure 2

Here we arrange the four figures into a multi-panel figure. 
```{r, fig.height=20, fig.width=14, "multipanel-fig"}

# With a lot of grid-foo to align the panels.

###
SingleFunc <- SingleFunc + theme(plot.margin=unit(c(1.5,0.5,1,1), "cm"))
SingleFunc_comp <- SingleFunc_comp + theme(plot.margin=unit(c(1.5,1,1,0.5), "cm"))
AverageFunc <- AverageFunc  + theme(plot.margin=unit(c(1,0.5,1,1), "cm"))
AverageFunc_comp <- AverageFunc_comp  + theme(plot.margin=unit(c(1,1,1,0.5), "cm"))
singleThresh <- singleThresh + theme(plot.margin=unit(c(1,0.5,1,1), "cm"))
singleThresh_comp <- singleThresh_comp + theme(plot.margin=unit(c(1,1,1,0.5), "cm"))
multiTresh <- multiTresh + theme(plot.margin=unit(c(1,0.5,1,1), "cm"))
multiTresh_comp <- multiTresh_comp + theme(plot.margin=unit(c(1,1,1,0.5), "cm"))


# convert plot to gtables
SingleFunc <- ggplotGrob(SingleFunc)
SingleFunc_comp <- ggplotGrob(SingleFunc_comp)
AverageFunc <- ggplotGrob(AverageFunc)
AverageFunc_comp <- ggplotGrob(AverageFunc_comp)
singleThresh <- ggplotGrob(singleThresh)
singleThresh_comp <- ggplotGrob(singleThresh_comp)
multiTresh <- ggplotGrob(multiTresh)
multiTresh_comp <- ggplotGrob(multiTresh_comp)

# convert width to unit lists 
SingleFunc$widths <- grid:::unit.list(SingleFunc$widths)   
SingleFunc_comp$widths <-  grid:::unit.list(SingleFunc_comp$widths)
AverageFunc$widths <- grid:::unit.list(AverageFunc$widths)   
AverageFunc_comp$widths <- grid:::unit.list(AverageFunc_comp$widths)
singleThresh$widths <- grid:::unit.list(singleThresh$widths)
singleThresh_comp$widths <- grid:::unit.list(singleThresh_comp$widths)   
multiTresh$widths <- grid:::unit.list(multiTresh$widths)   
multiTresh_comp$widths <- grid:::unit.list(multiTresh_comp$widths)   

# extract the first three widths, corresponding to left margin, y lab, and y axis
SingleFunc.widths <- SingleFunc$widths[1:3] 
SingleFunc_comp.widths <- SingleFunc_comp$widths[1:3] 
AverageFunc.widths <- AverageFunc$widths[1:3] 
AverageFunc_comp.widths <- AverageFunc_comp$widths[1:3] 
singleThresh.widths <- singleThresh$widths[1:3] 
singleThresh_comp.widths <- singleThresh_comp$widths[1:3] 
multiTresh.widths <- multiTresh$widths[1:3] 
multiTresh_comp.widths <- multiTresh_comp$widths[1:3] 

# extracting max width
max.widths <- unit.pmax(SingleFunc.widths, 
                        SingleFunc_comp.widths,
                        AverageFunc.widths,
                        AverageFunc_comp.widths,
                        singleThresh.widths,
                        singleThresh_comp.widths,
                        multiTresh.widths,
                        multiTresh_comp.widths)

# assigning max width
SingleFunc$widths[1:3] <-  max.widths
SingleFunc_comp$widths[1:3] <-  max.widths
AverageFunc$widths[1:3] <-  max.widths
AverageFunc_comp$widths[1:3] <-  max.widths
singleThresh$widths[1:3] <-  max.widths
singleThresh_comp$widths[1:3] <-  max.widths
multiTresh$widths[1:3] <-  max.widths
multiTresh_comp$widths[1:3] <-  max.widths
####

Figure_2 <- plot_grid(SingleFunc, SingleFunc_comp,
          AverageFunc, AverageFunc_comp,
          singleThresh, singleThresh_comp,
          multiTresh, multiTresh_comp, 
          ncol = 2,
          labels = letters[c(1, 5, 2, 6, 3, 7, 4, 8)],
          hjust = -6,
          vjust = 1.5,
          label_size = 18)

ggsave( "Figure_2.pdf", Figure_2, height = 20, width = 14, device=cairo_pdf)

Figure_2

```
